Proposed Solution

### 2017-05-05 ISSTA17/response.txt

We thank the reviewers for their time and positive feedback.

@reviewerA: "What are the core contributions to an ICFP audience?"

We agree with the reviewer that the translation from automata to FRP
is mainly a technical result. However, let us try to convince you that
this translation is just one of several crucial building blocks needed
for our major contribution: the first synthesis approach for
FRP. Specifically,

1. we introduce TSL, the first temporal logic for the specification of
   data transformations over time,

2. we formulate the technical constructions needed to translate TSL to
   LTL, and automata to FRP, and

3. we show that our approach works on examples that are significantly
   out of reach for existing reactive synthesis methods.

Overall, our paper establishes the first connection between the
research areas of FRP and reactive synthesis, and shows that this
connection provides significant value to both sides.

We also agree with the reviewer that, in synthesis, scalability is
always a serious concern. The translation of the automaton to an FRP
program is, however, very much not the bottleneck. The two aspects
that are critical to scalability are the following:

1. The synthesis of the automaton from the LTL formula. This is a
   double-exponentially hard problem, which, as we know from the
   reactive synthesis community, cannot be applied to arbitrarily
   large LTL formulas. We do not consider this as a serious issue for
   two reasons: a) there are effective methods developed in the
   reactive synthesis community which we leverage, such as bounded and
   symbolic synthesis techniques, and b) we are in much better shape
   than the reactive synthesis community, because we only synthesize
   the control structure (which tends to be reasonably small), while
   they also synthesize the data transformation (which tends to be
   arbitrarily large).

2. The under-approximation, which is necessary because the TSL
   synthesis problem is undecidable. We have not found this to be a
   problem in our experiments. If needed, the under-approximation can
   be improved manually.

@reviewerC: "Why should TSL be prefered over LTL?"

TSL is fundamentally more expressive than LTL. While LTL only allows
reasoning about Boolean data streams, TSL describes streams of
abstract data types. TSL is also a much more appropriate logic to
describe FRP programs, which typically handle complex data. The power
of TSL is also reflected in that its synthesis problem is in general
undecidable. We show that, by separating control and data, we can use
LTL synthesis algorithms to synthesize programs from TSL
specifications, which has proven effective on substantial FRP
programs.

@reviewerC: "I would have expected infinite state transition systems."

Yes and no. Our method separates control and data. While the control
part is (as usual) finite, the data part can be infinite. In that
sense, the reviewer is correct in observing that our reasoning is
finite by using uninterpreted functions, which provide a finite
abstraction.

@reviewerB:

We thank you for your encouraging words, and we like your suggestions
on how to restructure the paper to improve it further.

### 2017-05-05 ISSTA17/secs/completeConfigC.tex
expaneded version of FMCAD for ISSTA

This algorithm is complete but unsound.
We first define the set of incorrect files based on ground truth $Inc_{gt}$, and the set of files which are predicted by the model to be incorrect, $Inc_{prd}$.
To define $Inc_{gt}$, we will introduce the set of relations that are necessary and breaking based on ground truth, $Nec_{gt}$ and $Br_{gt}$ respectivly.
The definition of $Inc_{prd}$ follows from the description of the algorithm.

@@

To show completness we show that the $Inc_{gt} \subseteq Inc_{prd}$.

\begin{align*}
Nec \subseteq Nec_{gt}
%Nec \disjoint Br_{gt}
%\forall C \in {Incorrect(C)}, C \in system(Inc,C)   Nec_{True}, r \in Nec_{Learned} - (Nec_{Learned} \contains Nec_{True})
%\not \exists C \in SysInc(C), C \notin {Incorrect(C)} 
%\Nec_{Learned} \contains Nec_{Observable}
%\forall r \in Br, r \notin Incorrect(C) \implies \exists r \in Nec_{True}, r \notin M(C)
%Incorrect(C) \implies \exists r \in Nec, r \notin M(C)
\end{align*}

%The algorithm is complete if, for any file which breaks a relation for which we have a template (a misconfiguration) and have seen an correct example, the system will mark that file as incorrect.
%Since the model always maintins the strongest conditions for correctness for any file, the algorithm is clearly complete.
%Completness cooresponds to no false negatives.
%This algorithm is however unsound, some correct file may be marked as incorrect, a false postive.
%We have a high false postive rate since the model is an overappoximation of the true model.

>>>>

ConfigC is complete but unsound.
We first define the set of incorrect files based on ground truth $Inc_{gt}$, and the set of files which are predicted by the model to be incorrect, $Inc_{prd}$.
To define $Inc_{gt}$, we will introduce the set of relations that are necessary and breaking based on ground truth, $Nec_{gt}$ and $Br_{gt}$ respectively.
The definition of $Inc_{prd}$ follows from the description of the algorithm.

@@

Completeness follows from showing that $Inc_{gt} \subseteq Inc_{prd}$.
The algorithm is complete if, for any file which breaks a relation for which we have a template (a misconfiguration) and have seen an correct example, the system will mark that file as incorrect.
Since the model always maintains the strongest conditions for correctness for any file, the algorithm is clearly complete.
Completeness corresponds to no false negatives.
However, this algorithm is unsound, as some correct file may be marked as incorrect, i.e. ConfigC returns a false positive.
We have a high false positive rate since the model is an over-approximation of the true model.

### 2017-05-05 ISSTA17/secs/intro.tex
expaneded version of FMCAD for ISSTA

Machine learning has been used in various software analysis techniques, such as programming-by-example \cite{lau2000version}, invariant synthesis \cite{garg2014ice}, and error detection \cite{Santolucito2016}.
However, many popular machine learning algorithms, such as neural nets and n-gram models, are not designed to provide simple justifications for their classification results.
While effective in practice, the lack of justification for the results limits the applicability to software analysis, where justification is often critical.
For example, in the case of error detection, the system should not only report which files have errors, but also locate the errors.
Version space learning is a technique for logical constraint classification~\cite{mitchell82}.
As a logical constraint system, version space learning can provide the justification that is difficult to obtain from other methods, but its main weakness is that it cannot handle noisy data.

%However, many popular machine learning algorithms, such as neural nets and n-gram models, produce probabilistic models of correctness.
%While effective in practice, these approachs cannot be garunteed to be complete, that is they will always find the error.
%We use version space learning to address the need for a formal completness garuntee in automated model generation for verification.

In our previous work~\cite{Santolucito2016} we built a prototype to verify configuration files from a set of correct examples.
The tool, ConfigC, takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if a new configuration file adhears to those rules.
If a file is incorrect, the tool identifies the location of error, and reports what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so" before "extension recode.so"}.
While this prototype's results are promising, the main weakness are that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

In order to extend the prototype, in this paper we provide a new description of ConfigC in terms of version space learning.
This description will allow us to understand how the justifications are produced in ConfigC.
We can then extend ConfigC with a new algorithm to decrease the false positive rate, while maintaining the ability to provide justification.
%and show that the results are garunteed to be complete.

%Although the results are complete, the approach used previously produces a high false positive rate.
We propose an extension to ConfigC that allows it to handle both correct and incorrect files in the training set.
Additionally, we extend ConfigC with the ability to use structure within the training set.
While the previous training set from ConfigC was a random sampling of configuration files, many other training sets have a structure based on version history.
%The algorithm we propose takes advantage of this commonly availible, but underutilized structure of training sets for software analysis.
%Because most code does not exist in isolation, but changes over time with development, any training set of code from a verison control system, like Github, has a rich temporal structure. 
This structure is simply a partial order, and can be used by our proposed algorithm for more effective learning.
%A learning algorithm that uses the temporal structure of code to create sound classification with a low false positive rate can be used in at least the above listed software analysis techniques.

To test this approach in practice, we plan to implement our algorithm to check for TravisCI\footnote{\url{http://www.travis-ci.com}} configuration errors.
TravisCI is a continuous integration tool connected to Github that allows programmers to automatically run their test suite on every code update (commit).
A user adds a configuration file to the repository that enables TravisCI and specifies build conditions, such as which compiler to use, which dependencies are required, and a set of benchmarks to test.
This ensures the tool can always be automatically built correctly on a fresh machine.

@@

Our main contributions are then as follow:

@@

\item We give a description of our prototype tool, ConfigC, in the context of version space learning.
\item We propose a new algorithm for ConfigC to handle both incorrect and correct training data, as well as internal structure of a training set.
\item We propose a real-world application of this approach, and outline the steps needed for an implementation.

>>>>

%what is the problem
Configuration files provide easy access to the most critical parameters of software systems in order to quickly tune the behavior system.
This expressive power in a single file creates a large surface for potential errors that can degrade performance or take down systems entirely.
The verification task for configuration files cannot be addressed with traditional software analysis techniques.
There are rarely formal specifications of a correct configuration, 
  the source code of the underlying system may not be available, 
  and the configuration may depend on the external environment such as the network or available hardware.
To address this issue, we propose a method for automated configuration file verification using a data set of labeled examples of correct and incorrect configurations.

%what approach do we take to the problem
In previous work, we used version space learning for configuration file verification.
We extend that work which only works on correct examples, to work on temporally structured correct and incorrect examples by building an SMT formula to find a classification model.
As a demonstration of this algorithm, we plan to implement it to automatically verify TravisCI configurations.
The algorithm will utilize Github histories to allow us to detect potential build errors without actually building, saving valuable programmer and server time.

%why cant we use existing approach
%legibility issue
Machine learning has been used in various software analysis techniques, such as programming-by-example \cite{lau2000version}, invariant synthesis \cite{garg2014ice}, and error detection \cite{Santolucito2016}.
However, many popular machine learning algorithms, such as neural nets and n-gram models, are not designed to provide simple justifications for their classification results.
While effective in practice, the lack of justification for the results limits the applicability to software analysis, where justification is often critical.
For example, in the case of error detection, the system should not only report which files have errors, but also locate the errors.
This justification issue is also called \textit{legibility} of machine learning.

%correctness issue
Additionally, these probabilistic approaches to machine learning cannot be guaranteed to be complete, that is they will always find the error.
We use version space learning to address the need for a formal completeness guarantee in an automated model generation for verification.
Version space learning is a technique for logical constraint classification~\cite{mitchell82}.
As a logical constraint system, version space learning can provide the justification that is difficult to obtain from other methods, but its main weakness is that it cannot handle noisy data.

%What have we already done (labeled correct)
In our previous work~\cite{Santolucito2016} we built a prototype to verify configuration files from a set of correct examples.
The tool, ConfigC, takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if a new configuration file adheres to those rules.
If a file is incorrect, the tool identifies the location of error and reports what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so"} before \texttt{"extension recode.so"}.
While this prototype's results are promising, the main weakness is that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

%What do we 'plan' to do to extend this (labeled correct and incorrect)
In order to extend ConfigC, in this paper we provide a new description of ConfigC in terms of version space learning.
This description will allow us to understand how the justifications are produced in ConfigC.
We can then extend ConfigC with a new algorithm to decrease the false positive rate, while maintaining the ability to provide justification.

We propose an extension to ConfigC that allows it to handle both correct and incorrect files in the training set.
Additionally, we extend ConfigC with the ability to use structure within the training set.
While the previous training set from ConfigC was a random sampling of configuration files, many other training sets have a structure based on version history.
The algorithm we propose takes advantage of this commonly available, but underutilized structure of training sets for software analysis.
Because most code does not exist in isolation, but changes over time with development, any training set of code from a version control system, like Github, has this rich temporal structure. 
This structure is a partial order over time, and can be used by our proposed algorithm for more effective learning.

%evaluation plans
To test this approach in practice, we plan to implement our algorithm to check for TravisCI\footnote{\url{http://www.travis-ci.com}} configuration errors.
TravisCI is a continuous integration tool connected to Github that allows programmers to automatically run their test suite on every code update (commit).
A user adds a configuration file to the repository that enables TravisCI and specifies build conditions, such as which compiler to use, which dependencies are required, and a set of benchmarks to test.
This ensures the tool can always be automatically built correctly on a fresh machine.

@@

Our main contributions are as follow:

@@

\item We give a description of our tool, ConfigC, in the context of version space learning.
\item We propose a new algorithm to handle both incorrect and correct training data, as well as the internal structure of a training set.
\item We propose a real-world application of this approach, and outline the steps needed for an implementation.

### 2017-05-05 ISSTA17/secs/related.tex
expaneded version of FMCAD for ISSTA

\para{Language-support misconfiguration checking}
There have been several language-support efforts proposed for preventing
configuration errors introduced by fundamental deficiencies in
either untyped or low-level languages. For example, in the network
configuration management area, administrators often
produce configuration errors in their routing configuration files.
PRESTO~\cite{enck07configuration} 
automates the generation of device-native configurations
with configlets in a template language. 
Loo {\em et al.}~\cite{loo05declarative} adopt Datalog to reason about 
routing protocols in a declarative fashion. 
COOLAID~\cite{chen10declarative} constructs
a language to describe domain knowledge about devices and
services for convenient network reasoning and management.
Compared with the above efforts, \app mainly focuses on software systems, 
\eg, MySQL and Apache, and our main purpose is to automate configuration
verification rather than proposing new languages 
to convenient configuration-file writing. 
The closest effort to \app is ConfigC~\cite{santolucitoCAV},
which aims to learn configuration-checking rules from a given training
set. Compared with \app, ConfigC has the following disadvantages.
First, ConfigC requires the configuration files in the training set must 
be correct, which is impractical because it is very difficult to
determine a correct configuration set in reality.
Second, ConfigC covers fewer types of misconfigurations than \app. 
Finally, the training time of ConfigC is much longer than \app.

\para{Misconfiguration detection}
Misconfiguration detection techniques aim at checking configuration
efforts before system outages occur.
Most existing detection approaches check 
the configuration files against a set of predefined correctness 
rules, named constraints, and then report errors if 
the checked configuration files do not satisfy these rules.
Huang {\em et al.}~\cite{huang15confvalley},
for example, proposed a 
language, ConfValley, to validate 
whether given configuration files meet administrators' specifications. 
Different from \app, ConfValley does not
have inherent misconfiguration checking capability, since it only offers
a language representation and requires administrators to
manually write specifications, which is an error-prone
process. On the contrary, \app does not need users to manually
write anything.

@@

\para{Association Rule Learning}
The approach we have presented not only generalizes association rule learning, but also another learning strategy called \textit{sequential pattern mining}~\cite{mabroukeh2010taxonomy}. Ordering rules are similar to the patterns learned in sequential pattern mining, although we restrict our ordering rules to $|S|,|T|=1$ since these are the most common misconfigurations we encounter in practice.
While a major limitation to sequential pattern mining is the scalability of the problem~\cite{ayres2002sequential}, we escape this issue with $|S|,|T|=1$ and the fact that a single configuration file tends to be less a few thousand lines.
There has been other work in these same classes of algorithms~\cite{han2007frequent,langley1995applications} for various applications and variations on the core problem.
A future direction for this work is to integrate advances in these domains into the configuration file verification problem.

>>>>

\para{Language-support misconfiguration checking}
There have been several language-support efforts proposed for preventing
configuration errors introduced by fundamental deficiencies in
either untyped or low-level languages. For example, in the network
configuration management area, administrators often
produce configuration errors in their routing configuration files.
PRESTO~\cite{enck07configuration} 
automates the generation of device-native configurations
with configlets in a template language. 
Loo {\em et al.}~\cite{loo05declarative} adopt Datalog to reason about 
routing protocols in a declarative fashion. 
COOLAID~\cite{chen10declarative} constructs
a language to describe domain knowledge about devices and
services for convenient network reasoning and management.
Compared with the above efforts, our work focuses on software systems, 
\eg, MySQL and Apache, and our main purpose is to automate configuration
verification rather than proposing new languages 
to convenient configuration-file writing.

\para{Misconfiguration detection}
Misconfiguration detection techniques aim at checking configuration
efforts before system outages occur.
Most existing detection approaches check 
the configuration files against a set of predefined correctness 
rules, named constraints, and then report errors if 
the checked configuration files do not satisfy these rules.
Huang {\em et al.}~\cite{huang15confvalley},
for example, proposed a 
language, ConfValley, to validate 
whether given configuration files meet administrators' specifications. 
As opposed to our work, ConfValley does not
have inherent misconfiguration checking capability, since it only offers
a language representation and requires administrators to
manually write specifications, which is an error-prone
process. On the contrary, our work does not need users to manually
write anything.

@@

### 2017-05-05 ISSTA17/secs/temporal.tex
expaneded version of FMCAD for ISSTA

Since a TravisCI configuration file is dependent on the code it is trying to build, we must consider a more general sense of configuration file.
We will call this a program summary $P_t$, which is a representation of the repository which contains the information relevant to the learning process.
The subscript on $P_t$ is a timestamp tag based on the ordered git commit history.
In the case of TravisCI, this include the \verb|.travis.yml| file, as well as code features that may effect build status, such as programming language and a list of imported libraries.
The summary must be \textit{sufficiently detailed}, that is it must contain every piece of information that might lead to a build error.

%IMPORTANT, but no space
However, a git history is not a limited to a single linear timeline.
Git features the ability to \textit{branch}, which allows to simultaneous commit chains.
To handle the start of a branch, add a superscript to indicate the branch, and restart the counter on a branch.
To handle the merge of two branches $P_{t}^{x}$ and $P_{t'}^{y}$, step to $P_{t+1}^{x}$, where $x$ is the mainline branch.
We then say that $P_{t'}^{y}$ has no successor commit $P_{t'+1}^{y}$.
Note the ability to handle branching implies this method is suitable to any learning set which has a partial order.

@@

From the above we know that if a build is erroring, then there must exist at least one error.
%By pushing the negation into the formula,
We can also know that if a build is passing, then there must not exist any errors.
That is, the model of a passing commit must not contain any rules which are breaking.
Note we are not, however, guaranteed that any rules from a passing commit are necessary.
\begin{align}
  S(P_t) = Err \implies \exists r \in  M (P_t), r \in Br \label{eq:E}\\
  S(P_t) = Pass \implies \forall r \in  M (P_t), r \notin Br \label{eq:P}
\end{align}

%While Eq. \ref{eq:E} and \ref{eq:P} might build a basic model, they will do not capture all of the available knowledge.
%The key insight is that 
Additionally, when we commit a break ($PE$), we can localize the error to one of the relations that changed.
Either we removed something that was necessary, or added something that was breaking.
Note that this is an inclusive disjunction, since a erroring commit can break multiple things at once.
Expressed formally, where $\setminus$ is the set difference, that is:
\begin{align}
  S(P_{t,t+1}) &= PE \implies \nonumber \\
  \exists& r \in (M(P_{t})\ \setminus M(P_{t+1})), r \in Nec\ \nonumber \\
  \lor \ \exists& r \in (M(P_{t+1}) \setminus M(P_{t})), r \in Br \label{eq:PE}
\end{align}

We then can combine Eq \ref{eq:E}, \ref{eq:P}, and \ref{eq:PE} with conjunctions and ask a SMT solver for a model satifying the formula.
The resulting model will be the sets $Nec$, and $Br$, which can be used to check new configuration files.
Since we used a similar model to ConfigC, we will still be able to provide justifications for the classification results.

>>>>

While in ConfigC, configuration files were labeled in isolation from their system by an expert, TravisCI configurations are automatically labeled with respect to their containing repository.
Because the correctness of TravisCI configuration files is dependent on the code it builds, we must model more than just the configuration file, $C$. 
We introduce a program summary $P_t$, which is a combination of $C$ as well as the information relevant to the learning process from the repository.
The subscript on $P_t$ is a timestamp tag based on the ordered git commit history.
In the case of TravisCI, this include the \verb|.travis.yml| file, as well as code features that may effect build status, such as programming language and a list of imported libraries.
The summary must be \textit{sufficiently detailed}, that is it must contain every piece of information that might lead to a build error.

%IMPORTANT, but no space
However, most versioning systems do not limit version history to a single linear timeline.
For example, Git features the ability to \textit{branch}, which allows to simultaneous commit chains.
To handle the start of a branch, add a superscript to indicate the branch, and restart the counter on a branch.
To handle the merge of two branches $P_{t}^{x}$ and $P_{t'}^{y}$, step to $P_{t+1}^{x}$, where $x$ is the mainline branch.
We then say that $P_{t'}^{y}$ has no successor commit $P_{t'+1}^{y}$.
Note the ability to handle branching implies this method is suitable to any training set which has a partial order.

@@

From the above we know that if a build is erroring, then there must exist at least one error.
We can also know that if a build is passing, then there must not exist any errors.
That is, the model of a passing commit must not contain any rules which are breaking.
Note we are not, however, guaranteed that any rules from a passing commit are necessary.
\begin{align}
  S(P_t) = Err \implies \exists r \in  M (P_t), r \in Br \label{eq:E}\\
  S(P_t) = Pass \implies \forall r \in  M (P_t), r \notin Br \label{eq:P}
\end{align}

%While Eq. \ref{eq:E} and \ref{eq:P} might build a basic model, they will do not capture all of the available knowledge.
%The key insight is that 
Additionally, when we commit a break ($PE$), we can localize the error to one of the relations that changed.
Either we removed something that was necessary, or added something that was breaking.
Note that this is an inclusive disjunction, since a erroring commit can break multiple things at once.
When comparing multiple versions of configurations, it is important that we look at the difference between the models $M(P)$ and not just $P$.
In the latter case, we would treat individual lines as error sources, when in fact we want to detect the relations between configuration settings.
Expressed formally, where $\setminus$ is the set difference, we learn the following from a breaking commit:
\begin{align}
  S(P_{t,t+1}) &= PE \implies \nonumber \\
  \exists& r \in (M(P_{t})\ \setminus M(P_{t+1})), r \in Nec\ \nonumber \\
  \lor \ \exists& r \in (M(P_{t+1}) \setminus M(P_{t})), r \in Br \label{eq:PE}
\end{align}

We then can combine Eq \ref{eq:E}, \ref{eq:P}, and \ref{eq:PE} with conjunctions and ask a SMT solver for a model satisfying the formula.
The resulting model will be the sets $Nec$, and $Br$, which can be used to check new configuration files.
Since we used a similar model to ConfigC, we will still be able to provide justifications for the classification results.

### 2017-05-05 ISSTA17/secs/vspace_verify.tex
expaneded version of FMCAD for ISSTA

\begin{center}
\begin{tabular} {|c |c |c |}
 \hline
 Template & Input  & Relation \\
\hline
 ordering & (Line,Line) & Before $\vert$ After $\vert$ None \\
 \hline
 integer relation & (Line,Line)  & $< | == | >$  \\
\hline
 type & (Line,Line)  & String $\vert$ Int $\vert$ Filepath $\vert$ IP  \\
\hline
 missing entry & (Line,Line) & Required $\vert$ Not \\
\hline
\end{tabular}
\end{center}

@@

%When the system is run on a new file, one of the relations from the templates will not be satisfied.
%To provide an error message, the system only needs to report which relation is not satisfied.

>>>>

\begin{center}
\begin{tabular} {|c |c |c |}
 \hline
 Template & Arity & Relation \\
\hline
 ordering &            2 & Before $\vert$ After $\vert$ None \\
\hline
 coarse relation &     2  & $< | == | >$  \\
\hline
 fine-grain relation & 3  & $< | == | >$  \\
\hline
 type &                1  & String $\vert$ Int $\vert$ Boolean $\vert$ IP  \\
\hline
 missing entry &       2 & Required $\vert$ Not \\
\hline
\end{tabular}
\end{center}

@@

When the system is run on a new file, one of the relations from the templates will not be satisfied.
To provide an error message, the system only needs to report which relation is not satisfied.
